---
title: "Communicable, Maternal, Neonatal, and Nutritional Disease Burden (CMNN)"
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message = FALSE, warning = FALSE)
```


```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the communicable diseases data
data_cmnn <- read_csv(here("data", "burden-of-disease-cmnn.csv"))

data_cmnn_2 <- data_cmnn %>% 
  filter(Entity %in% c("Angola", "Belgium", "Benin"))

```

## Table of Estimates for CMNN Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
# Use kable() from the knitr package to render the table.

# Pivot the data to show each country in a separate column

data_cmnn_3 <- data_cmnn_2 %>%
  rename(daly =`DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)`)

data_pivoted <- pivot_wider(data_cmnn_3,
                            id_cols = Year,
                            names_from = Entity,
                            values_from = daly)

# Render the pivoted table using kable
kable(data_pivoted)
# View(data_pivoted)

#?pivot_wider

```

## Plot Showing Trends in CMNN Burden Over Time

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.

plot <- ggplot(data = data_cmnn_3, aes(x = Year, y = daly, color = Entity)) +
  geom_line() +
  labs(x = "Year", y = "DALY", title = "Trend of DALY burden for the chosen countries over time") +
  scale_color_manual(values = c("Angola" = "red", "Belgium" = "blue", "Benin" = "green")) +
  theme_minimal()+
  scale_x_log10()
plot

```

## Summary of CMNN Burden Findings
This report analyzes the burden of CMNN conditions in three countries: Angola, Belgium, and Benin, over the period from 1990 to 2020.
Belgium exhibited the lowest Disability-Adjusted Life Years (DALY) burden among the three countries throughout the study period. This suggests relatively better health outcomes and healthcare infrastructure compared to Angola and Benin.






